---
title: "R_Workshop_Day2"
author: "Alexandria Jones"
date: "2024-02-20"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# This is a chunk in Rmd
```{r}
plot(1:10)
```

###################################
# 1.2 Get to know the data
###################################
# Read in the data
```{r}
# Read in the two files
dat_1 = read.csv("RNAseq_Data_1.csv")
dat_2 = read.csv("RNAseq_Data_2.csv")

# Look at the head() of each data set
head(dat_1)
head(dat_2)
```

# Challenge Question
# What do you observe when you look at dat_1?
# i observe a data table with a bunch of information about multiple genes.
# What do you observe when you look at dat_2?
# i observe another data table with lots of info on different genes.

# Let's look at the dimensions and column names of dat_1
```{r}
# This function tells you the number of rows and columns
dim(dat_1)

# The function prints the names of each of the columns
colnames(dat_1)
```

###################################
# 2.3 Try out different tables
##################################
# Summary function
```{r}
# Summarize Sample BC_1_RPKM in dat_1
summary(dat_1$BC_1_fpkm)

# Summarize the Gene Length in dat_1
summary(dat_1$GeneLength)
```

# Table
```{r}
# Count the number of chromosomes are represented in the data
table(dat_1$Chrom)

# Count the number of gene types
table(dat_1$GeneType)
```
# Summarise
```{r}
# Load the appropriate packages that are not built into R
library(dplyr)

# Summarize the gene length by chromosome
summary_data1 <- dat_1 %>%
  group_by(Chrom) %>%
  summarise(mean_geneLength = mean(GeneLength), sum_geneLength = sum(GeneLength))

summary_data1


# Summarize the gene length by GeneType
summary_data2 <- dat_1 %>%
  group_by(GeneType) %>%
  summarise(mean_geneLength = mean(GeneLength), sum_geneLength = sum(GeneLength))

summary_data2
```
# Summarize the gene length by Chromosome for dat_1
# Calculate the average gene length, total gene length, median gene length, and max gene length
```{r}
summary_data1 <- dat_1 %>%
  group_by(Chrom) %>%
  summarise(mean_geneLength = mean(GeneLength), sum_geneLength = sum(GeneLength), median_geneLength = median(GeneLength), max_geneLength = max(GeneLength))
summary
```


# Join both dat_1 and dat_2 tables together
```{r}
# Join data using left_join()
dat_joined_left = left_join(dat_1, dat_2, by = c("geneID","GeneName","Chrom","Strand", "Start","End","GeneLength", "GeneType", "GeneDescription"))

# Join data using inner_join()
dat_joined_inner = inner_join(dat_1, dat_2, by = c("geneID","GeneName","Chrom","Strand", "Start","End","GeneLength", "GeneType", "GeneDescription"))

# Look at the column names of the data dat_joined_inner
colnames(dat_joined_inner)
# Reorganize data so GeneType and GeneDescripton is at the end of the table
dat_joined_inner = dat_joined_inner[,c(1:13,16:21,14:15)]

head(dat_joined_inner)
```
# Put it all together
```{r}
# Make a new dataset
dat_new = dat_joined_inner

# Calculate the mean of experiment BC
dat_new$BC_fpkm_average = rowMeans(dat_new[,c(8:10)])

# Calculate the standard deviation of BC using the apply() function
dat_new$BC_fpkm_sd = apply(dat_new[,c(8:10)], 1, sd)

# Calculate the mean of experiment BC_DOX
dat_new$BC_DOX_fpkm_mean = rowMeans(dat_new[,c(11:13)])
# Calculate the standard deviation of BC_DOX
dat_new$BC_DOX_fpkm_sd = rowMeans(dat_new[,c(11:13)])
# Calculate the mean of experiment DL
dat_new$DL_fpkm_mean = rowMeans(dat_new[,c(14:16)])
# Calculate the standard deviation of DL
dat_new$DL_fpkm_sd = rowMeans(dat_new[,c(14:16)])
# Calculate the mean of experiment DL_DOX
dat_new$DL_DOX_fpkm_mean = rowMeans(dat_new[,c(17:19)])
# Calculate the standard deviation of DL_DOX
dat_new$DL_DOX_fpkm_sd = rowMeans(dat_new[,c(17:19)])
head(dat_new)
```

# Now save your data to a csv file and download it to your files
```{r}
write.csv(dat_new, "my_analysis.csv")
```

###################################
# STEP 3 GGPLOT
###################################

# 3.1 GGplot Training challenge
# 3.2 GGplot more in-depth walk-though
#######################################
# 3.2 GGplot more in-depth walk-though
#######################################

```{r}
# Load the tidyverse library
library(tidyverse)
# Load the ggplot2 library
library(ggplot2)

# Create a sample data frame
data <- data.frame(
  Group = rep(c("A", "B", "C", "D"), each = 20),
  Value = rnorm(80)
)

# Create the boxplot
ggplot(data, aes(x = Group, y = Value)) +
  geom_boxplot() +
  xlab("Group") +
  ylab("Value") +
  ggtitle("Boxplot of Values by Group")
```

# Get your data for the Genes_FPKM.csv file into R
```{r}
# Read in the Genes_FPKM.csv file and assign it to the object Genes
Genes = read.csv("Genes_FPKM.csv")
# What are the number of rows and columns in this dataset?
dim(Genes)
# What type of data is this (matrix or dataframe)?
class(Genes)
# this is a dataframe
# Look at the top 6 rows of the dataset
head(Genes)

```
# Challenge question
# What do you think this data is showing?


# Create GAPDH boxplot
```{r}
# Load the ggplot2 library
library(ggplot2)

# Add color pallete
install.packages('palettetown')
library(palettetown)
# https://github.com/timcdlucas/palettetown

# Make colors!
my_colors <- pokepal(pokemon = 137, spread = 4)


# Create the boxplot
ggplot(Genes, aes(x = Experiment.Group, y = GAPDH_FPKM, fill = Experiment.Group)) +
  geom_boxplot() +
  xlab("Group") +
  ylab("GAPDH_FPKM") +
  scale_fill_manual(values = my_colors) +
  ggtitle("FPKM of GAPDH")

```
# Challenge question
# What do you think each element of the ggplot code does?
# each element of the ggplot code coordinates between the x and y plots of data.
# ggplot(Genes, aes(x = Experiment.Group, y = GAPDH_FPKM, fill = Experiment.Group)) +

#  geom_boxplot() +

#  xlab("Group") +

#  ylab("GAPDH_FPKM") +

#  scale_fill_manual(values = my_colors) +

#  ggtitle("FPKM of GAPDH")


# Create H19 boxplot
# Be sure to include appropriate comments and code
# Use different colors than for the GAPDH plot
```{r}
my_colors <- c("pink","black","purple","grey")

ggplot(Genes, aes(x = Experiment.Group, y = H19_FPKM, fill = Experiment.Group)) +
  geom_boxplot() +
  xlab("Group") +
  ylab("H19_FPKM") +
  scale_fill_manual(values = my_colors) +
  ggtitle("FPKM of H19")


```

# Challenge question
# Describe what each plot is showing
# each plot is showing the H19 FPKM data.
# Compare the two plots to eachother, what is similar and what is different?
# the sizes of the boxes are different, and so is the data.